---
title: "DATA 607 - Project 1"
author: "Emrah Akin"
date: "9/24/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r results='hide', message=FALSE, warning=FALSE, echo=F}
library(stringr)
```

#### In this project, you’re given a text file with chess tournament results where the information has some structure. Your job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database) with the following information for all of the players:

#### Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents

<br>

```{r}
url <- "https://raw.githubusercontent.com/emrahakin1985/DATA607/master/datasets/tournamentinfo.txt"
chess_x <- read.csv(url, stringsAsFactors = F, skip = 3)

#how regex sees it:
head(unlist(str_extract_all(chess_x, ".+")), 2)

```


```{r}
chess <- unlist(chess_x)
chess_reg <- unlist(str_extract_all(chess, ".+"))
# how regex sees it:
head(chess_reg, 11)
```

<br>

**First step is to narrow down, second step is to get the vector with full names only:**
```{r}
names_uncln <- unlist(str_extract_all(chess, "\\| +[A-Z]{2,}.+[A-Z]{2,} +\\|?"))

head(names_uncln, 10)

```

<br>

**Second step:**
```{r}
names <- unlist(str_extract_all(unlist(names_uncln), "[A-Z].+[A-Z]"))
head(names, 10)
```

<br>

**States:**
```{r}
state <- unlist(str_extract_all(chess, "^ +[A-Z]{2}"))
state <- str_trim(state)
#
state
#
```

<br>

**Points:**
```{r}
points <- unlist(str_extract_all(chess, "\\d\\.\\d"))
#
points
#
```

<br>

**Pre_rates:**
```{r}
pre_rate_uncln <- unlist(str_extract_all(chess, "R\\: +\\d{3,4}P?(\\d{2})?"))
pre_rate_almost_done <- unlist(str_extract_all(unlist(pre_rate_uncln), "\\d.+\\d"))
pre_rate <- str_replace(unlist(pre_rate_almost_done), "P", ".")
#
pre_rate
#
```

<br>

**2 step regex code to extract opponent id's:**
```{r}
step1 <- "\\|[A-Z] {2,}\\d{1,2}"
step2 <- "\\d{1,2}"
```

<br>

**finding the rows where we will extract the data for opponent id's:**
```{r}
vector_test <- ""
for(i in 1:length(chess)){
  if (str_detect(chess[i], step1) == T){
    vector_test[i] <- i
  } 
}
vector_test <- vector_test[!is.na(vector_test)]
vector_test
```

```{r}
# creating data frame with established columns
final_df <- data.frame(names, state, points, pre_rate)
final_df[1:10,]
```

<br>

**Extracting and placing opponent id's for each player in a new column:**
```{r}

k <- 1
for(i in as.numeric(vector_test)){
  #print(i)
  final_df$Opponent_ID[k] <- str_c(unlist(str_extract_all(unlist(unlist(str_extract_all(chess[i], step1))), step2)), collapse = " ")
  k <- k + 1
}

final_df[c(1:5,55:59),]
```

<br>

**First, we need to change the type of pre_rate values (from factors to characters first, then numerical values) to make the calculations.**
```{r}
final_df$pre_rate <- as.numeric(as.character(final_df$pre_rate))
```

<br>

**Matching opponent id and their rate and calculating their average for each row (player):** 
```{r}
x <- 0

for( i in 1:nrow(final_df) ) {
  
  for ( u in 1:length(unlist(str_split(final_df$Opponent_ID[i], " ")  )  )   ){
    
    x <- x + final_df$pre_rate[as.numeric(   unlist(str_split(final_df$Opponent_ID[i], " "))[u]   )]
    #print(x)
    
  }
  final_df$Average_Opponent_Rate[i] <- x / length(unlist(str_split(final_df$Opponent_ID[i], " ")  )  )
  #print(x)
  x<-0
}
```

```{r}
final_df$Average_Opponent_Rate <- round(final_df$Average_Opponent_Rate,0)
final_df <- subset(final_df,,-5)
knitr::kable(final_df)
```

```{r}
write.csv(final_df, "chess.csv")
```