---
title: "DATA 607 Project 2a"
author: "Emrah Akin"
date: "10/8/2017"
output: html_document
---

```{r results='hide', message=FALSE, warning=FALSE, echo=F}
library(tidyverse)
library(stringr)
library(dplyr)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
url_drugs <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/drug-use-by-age/drug-use-by-age.csv"
drug <- read.csv(url_drugs, stringsAsFactors = F)
```

<br>

**Renaming some columns to make separating with "." easier:**
```{r}
drug2 <- drug %>%
  rename("pain releiver.use" = pain.releiver.use,
         "pain releiver.frequency" = pain.releiver.frequency)
knitr::kable(head(drug2, 10))

```

<br>

**Gathering columns, under key and values:**
```{r}
drug_gathered <- suppressWarnings(drug2 %>% 
  gather(key, value, -age, -n) %>% 
  arrange(age))
  

head(drug_gathered, 10)
```

<br>

**Separating key into subs and variables before spreading variables into 2 columns:**
```{r}
drug_gathered2 <- drug_gathered %>% 
  separate(key, into = c("subs", "variables"), sep = "\\." )
head(drug_gathered2,10)
```

<br>

**Spreading variables into use and frequency columns:**
```{r}
drug_gathered3 <- drug_gathered2 %>% 
  spread(variables, value)
  

knitr::kable(head(drug_gathered3, 15))

```

<br>

**Table  alcohol use:**
```{r}
drug_gathered3 <- suppressWarnings(drug_gathered3 %>% 
  mutate(use = as.numeric(use),
         frequency = as.numeric(frequency)))

alcohol <- drug_gathered3 %>% 
  filter(subs=="alcohol") %>% 
  select(-subs)

head(alcohol)

```

<br>

**Association between age and alcohol usage - frequency and age:**
```{r}
ggplot(alcohol) + geom_point(aes(age, y = use), color = "blue") + labs(x = "Age", y = "Median Use")
ggplot(alcohol) + geom_point(aes(age, y = frequency), color = "red") + labs(x = "Age", y = "Freuency")
```

<br>

**Table for 16 year old drug users:**
```{r}
age_16 <- drug_gathered3 %>% 
  filter(age == 16 & subs != "alcohol") %>% 
  select(-age) 
head(age_16)
```

<br>

**Bar plot for median drug use:**
```{r}
ggplot(age_16, aes(reorder(subs,-use),use), ylim = c(0,50) ) + 
  geom_bar(stat = "identity", color = "orange", fill = "white") +
  labs(x = "Median Drug Use", y = "Drugs") +
  coord_fixed(ratio = .2) +
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
```

