---
title: "607 - Week3 Assignment"
author: "Emrah Akin"
date: "9/14/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r results='hide', message=FALSE, warning=FALSE, echo=F}
library(stringr)
```
<br>
**1- Arranege the vector, "simp" so that all elements conform to the standard first name and last name.**
```{r}
simp <- "555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"
```
<br>
Extracting full names
```{r}
simpnames_orig <- unlist(str_extract_all(simp, "[[:alpha:]., ]{2,}"))
simpnames <- simpnames_orig
simpnames_orig
```
<br>
**Changing the format to first_name  last_name **
```{r}
for( i in 1:length(simpnames)){
  if(str_detect(simpnames[i], ",") == TRUE){
    simpnames[i] = str_c(str_replace(str_extract(simpnames[i], "(, ).+"), ", ", "")
                         , " ", str_replace(str_extract(simpnames[i], ".+,"), ",", ""))
  }else if(str_detect(simpnames[i], "[[:alpha:]]{2,}+\\. ") == T){
    simpnames[i] = str_replace(simpnames[i], str_extract(simpnames[i], "[[:alpha:]]{2,}+\\. "), "")
    }
}

simpnames
```
<br>
**2- Construct a logical vector indicating whther a character has a title (i.e., Rev. and Dr.).**

```{r}
str_detect(simpnames_orig, "^[[:alpha:]]{2,}+\\. ")
```
<br>
**3. Construct a logical vector indicating whther a character has a second name.**
```{r}
str_detect(simpnames_orig, "(^[[:alpha:]]{1}\\.) | [[:alpha:]]{1}\\.")
str_detect(simpnames, "(^[[:alpha:]]{1}\\.) | [[:alpha:]]{1}\\.")
```
<br>
**4- Describe the types of strings that conform to the following regular expressions and construct an example that is matched by regular expression.**
<br>
(a)[0-9]+\\$   At least one digit followed by the dollar sign:
```{r}
x <- "$ dollar sign 5$ 1000000$ 1,000$"
unlist(str_extract_all(x, "[0-9]+\\$"))
```
<br>
(b)\\b[a-z]{1,4}\\b A word with a sequence of at least 1 and at most 4 lowercase letters will be matched.
```{r}
w <- "France, fra.nce, one, two, three, fo.ur, fra,nce98, a test"
unlist(str_extract_all(w, "\\b[a-z]{1,4}\\b"))
```
<br>
(c) .*?\\.txt$ Strings ending with .txt and any characters that come before that in the string, if there is any, will be matched.
```{r}
f <- c("test.txt", "test test test$ .txt",   " .txt",   "test.txt.")
unlist(str_extract_all(f, ".*\\.txt$"))
```
<br>
(d) \\d{2}/\\d{2}/\\d{4}  digits in the following format will be matched: xx/xx/xxxx
```{r}
d <- "33/22/1111, 10-11-1991, 7/4/1776, 07/04/1776" 
unlist(str_extract_all(d, "\\d{2}/\\d{2}/\\d{4}"))
```
<br>
(e) <(.+?)>.+?</\\1>  at least one character in tag and again at least one character in the content area, then the same string with literal "/" in the beginning.
```{r}
t <- "test test<test>m</t> <html>test</html>"
unlist(str_extract_all(t, "<(.+?)>.+?</\\1>"))
```

