---
title: "Assignment2"
author: "Emrah Akin"
date: "9/7/2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r results='hide', message=FALSE, warning=FALSE, echo=F}
require(RMySQL)
require(lubridate)
require(reshape)
require(dplyr)

```

```{r}
db_movies <- dbConnect(MySQL(), 
        user = 'data607', password = 'pass', host = 'localhost', dbname = 'Movies')
```

```{r}
dbListTables(db_movies)
```

```{r}
sql_movies <- "select m.title, m.release_date, c.name, r.rating from movies m join critics c on m.movieid = c.movieid join ratings r on r.ratingid = c.ratingid;"
movies <- dbGetQuery(db_movies, sql_movies)
colnames(movies)[3] <- "critic_name"
movies

# same dataframe can also be read in R from the following github link:
# url <- "https://raw.githubusercontent.com/emrahakin1985/DATA607/master/datasets/movies.csv"
# movies <- subset(read.csv(url),,-1)
```

Pivot operation in R to pivot movie titles into columns so we have a column for each movie title and one row for each critic.
```{r}
# using reshape function, cast
movies_reshaped <- cast(movies, critic_name ~ title, value = 'rating')
```

```{r}
knitr::kable(movies_reshaped)
```

Arranging the order so that movies have the same alphabetical order (before extracting year information) as the reshaped data frame.
```{r}
movies <- movies %>%
  arrange(title)

# using lubridate function, year
year_vector <- unique(year(movies$release_date))
year_vector
```

Formatting vector before using it to change the column names on reshaped dataset
```{r}
year_vector <- paste('(', year_vector, ')', sep = "")
year_vector
```


Adding year information to the columns on movies_reshaped.
```{r}
colnames(movies_reshaped)[2:7] <- paste (colnames(movies_reshaped)[2:7], year_vector, sep = " ")
movies_reshaped
```

Better looking table with kable
```{r}
knitr::kable(movies_reshaped)
```

